\subsection{Fixed Resolution (FR) Policy Algorithm}
\label{sec:fixed-resolution-algo}

The fixed resolution (FR) policy algorithm adopts a simplistic strategy: retain observation time points at intervals of a user-specified factor $r$.
This procedure equates to naive downsampling via decimation \citep[p. 31]{crochiere1983multirate}.
We include discussion of this policy algorithm primarliy for completeness.

The procedures to decide eliminated time points during the update process and to enumerate retained observation times are trivial.
Pruning targets time points that are not multiples of the fixed resolution $r$ and enumeration traverses time points at an even stride $r$ until reaching record depth.
% Algorithms \ref{alg:fixed-resolution-algo-gen-drop-ranks} and \ref{alg:fixed-resolution-algo-enum-retained-ranks} provide these procedures formally.

Unsurprisingly, extant record size order of growth is $\mathcal{O}(n)$.
A simple justification can be given: record depth $n$ provides an upper bound for extant record size because, being strictly subtractive, pruned size cannot exceed record depth.
Likewise, the downsampling factor $r$ enforces an upper bound $r$ on gap size between retained observations.
Any larger gap size would require at least one time point multiple of $r$ to have been discarded.
% \input{thm/fixed-resolution-algo-space-complexity.tex}
% \input{thm/fixed-resolution-algo-uncertainty-bound.tex}

The FR policy algorithm provides stable absolute accuracy indefinitely.
Such an approach will be necessary for scenarios that can tolerate only certain observational uncertainties.
However, because extant record size grows linearly this policy would not suit applications expecting long observational duration, high-frequency observation, or storage limitation.
Figure \ref{fig:retention-policies} includes a time-lapse of the extant record under the FR policy algorithm.

% TODO fill in if time available
% \input{alg/fixed-resolution-algo-gen-drop-ranks.tex}
% \input{alg/fixed-resolution-algo-enum-retained-ranks.tex}
