\section{Streaming Curation Algorithms} \label{sec:annotation-algorithms}

\input{fig/retention-policies.tex}

This section collects specification and validation of five proposed streaming curation algorithms.
These algorithms operate online on a rolling stream of incoming observations to maintain a representative subsample of retained observations.
Streaming curation algorithms differ in the growth rate allowed for the curated observation collection and with respect to relative prioritization of retaining recent observations compared to older observations.

% For each surveyed algorithm, we present
% \begin{itemize}
% \item narration of retention policy algorithm strategy,
% \item retention policy algorithm extant record size order of growth (i.e., the scaling relationship between record depth and curated collection size),
% \item resolution guarantees (i.e., bounds on gap size between retained observations),
% \item instructions to enumerate time points retained within the curated subsample for any record depth,
% \item validation of policy algorithm self-consistency (i.e., that preceding record depths retain all observation time points required at later record depths),
% \item an algorithm for policy enactment (i.e., what observations to drop at each successive depth), and
% \item reflection of likely use case scenarios.
% \end{itemize}

We introduce the following five curation policy algorithms,
\begin{itemize}
\item Fixed-Resolution (FR) Policy Algorithm (Section \ref{sec:fixed-resolution-algo}),
\item Depth-Proportional Resolution (DPR) Policy Algorithm (Section \ref{sec:depth-proportional-resolution-algo}),
\item Recency-Proportional Resolution (RPR) Policy Algorithm (Section \ref{sec:recency-proportional-resolution-algo}),
\item Geometric Seqpence $n$th Root(GSNR) Policy Algorithm (Section \ref{sec:geom-seq-nth-root-algo}), and
\item Curbed Recency-Proportional Resolution (CRPR) Policy Algorithm (Section \ref{sec:curbed-recency-proportional-resolution-algo}).
\end{itemize}
The accomanying \texttt{hstrat} library provides reference implementations for all five.
Figure \ref{fig:retention-policies} compares retention patterns induced by each algorithm and recaps each policy algorithm's principal properties.
FR and DPR follow even curation prioritization while RPR, GSNR, and CRPR follow recency-proportional curation prioritization.
Recency-proportional techniques are of particular interest due to predictions from coalescence theory \citep{nordborgCoalescentTheory2019, berestyckiRecentProgressCoalescent2009}, which suggest that in most cases there will be more lineage divergences in recent time points.
Thus, prioritizing recent data should often improve the accuracy of subsequent data analysis. 
% should we justify recency-proporitonal here?
% ELD: Added it
Collection size grows the most aggressively under FR as $\mathcal{O}(n)$.
RPR reduces collection size growth to $\mathcal{O}(\log n)$.
The remaining algorithms enforce a fixed cap on curated collection size.
Note that GSNR and CRPR exhibit identical asymptotic properties.
We include both, as CRPR is an engineered extension of GSNR that improves the efficacy of available space usage during initial shallow record depth.

Appropriate algorithm choice will depend on use case scenario.
Relevant criteria to consider include
\begin{itemize}
  \item uncertainty and magnitude of upper bounds on record depth, if any,
  \item available storage capacity,
  \item relative importance of recent and ancient observations, and
  \item any hard record quality requirements (i.e., maximum acceptable gap size).
\end{itemize}

In a real-time scenario, record depth bounds would be considered in terms of upper bounds on chronological duration of record collection and the real-time sampling rate of observations.

\input{text/body/annotation-algorithms/fixed-resolution-algo.tex}

\input{text/body/annotation-algorithms/depth-proportional-resolution-algo.tex}

\input{text/body/annotation-algorithms/recency-proportional-resolution-algo.tex}

\input{text/body/annotation-algorithms/geom-seq-nth-root-algo.tex}

\input{text/body/annotation-algorithms/curbed-recency-proportional-resolution-algo.tex}
